---
author: "Jacob Bellmund"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

# Analysis Setup

## Defining Variables and Folders

First, we set up some global variables that will be used throughout the analyses. This includes the identifiers of the subjects to include in the analysis, specifics of the design such as the number of days and events per day and the regions of interest for representational similarity analysis.

```{r}

#-------- DEFINE GLOBAL VARIABLES ---------

# list of subject IDs(excluded: 58 and 68 because of bad memory performance and MRI acquisition problems)
subjects <- c("031", "032", "033", "034", "035", "036", "037", "038", "039", "054", 
              "055", "056", "057", "059", "061", "062", "063", "064", "065", "066", 
              "069", "070", "071", "072", "073", "074", "075", "076")
n_subs <- length(subjects)

# PVT scanning runs and blocks within each run
runs <- c("pre", "post")
n_runs <- length(runs)
n_blocks <- 10

# design parameters
n_days <- 4
n_events_day <- 5
n_pics <- n_days*n_events_day

# Main regions of interest for analyses
rois <- c("aHPC_lr", "alEC_lr")

# Regions of interest in MNI space
rois_ec = c("alEC_lr")
rois_hpc = c("aHPC_lr")
rois_mni <- c(rois_hpc, rois_ec)
roi_colors <- c("#dd4731", "#079cd6")

# Regions of interest to get from Freesurfer
rois_fs = c("hpc_lr", "ec_lr")
# numeric Freesurfer labels
#(https://surfer.nmr.mgh.harvard.edu/fswiki/FsTutorial/AnatomicalROI/FreeSurferColorLUT)
labels_fs = list(c(17,53), c(1006,2006)) 
roi_colors_fs <- wes_palette(n = length(rois_fs), name = "FantasticFox1")

# Define colors to use in plots
event_colors <- scico::scico(n=5, begin = 0, end = 0.6, palette = "bamako")
time_colors <- c("#26588E", "#33602D", "#C1AA6E") 

aHPC_colors <- c("#dd4731", # main
                 "#26588E", # within main, from scico::scico(n=5, begin = 0.1, end = 0.7, palette = "devon")[3], 
                 "#e31a1c", # within low
                 "#800026", # within high
                 "#A54440", # across main, from scico::scico(n=5, begin = 0.3, end = 0.9, palette = "lajolla")[3]
                 "#feb24c", # across low
                 "#fc4e2a") # across high
names(aHPC_colors) <- c("main", "within_main", "within_low", "within_high", "across_main", "across_low", "across_high")
alEC_colors <- c("#855C85", 
                 "#225ea8", 
                 "#1d91c0", 
                 "#0c2c84", 
                 "#7fcdbb", 
                 "#c7e9b4", 
                 "#41b6c4")
names(alEC_colors) <- c("main", "within_main", "within_low", "within_high", "across_main", "across_low", "across_high")
day_time_int_color <- "#F5DF4D"
time_within_across_color <- unname(alEC_colors["main"])
ultimate_gray <- "#939597"
```

### Folder Structure {-}

In a second step, we create some folders that will be used during the analysis. These are folders that contain raw data as well as folders into which processed data is written. CAVE: Some Markdown files still create their own folders.

```{r}
#-------- SET UP FOLDERS ---------
dirs <- c()

# directory with logs from picture viewing task
dirs$pvt_log_dir <- here("data", "behavior", "logs", "picture_viewing_task")

# directory with logs from day learning task
dirs$dlt_log_dir <- here("data", "behavior", "logs", "day_learning")

# directory with logs from the timeline task
dirs$timeline_dat_dir <- here("data", "behavior", "timeline")

# directories for freesurfer ROIs
dirs$rois_fs_dirs = here("data", "mri", "rois", rois_fs) 

# directories for final ROIs in analysis space
dirs$rois_ss_dirs <- here("data", "mri", "rois", rois, "samespace")

# directories for the MNI ROIs in analysis space
dirs$rois_mni_ss_dirs <- here("data", "mri", "rois", rois_mni, "samespace")

# directory where preprocessed data lies for each run and block
dirs$feat_dir <- here("data", "mri", "processed", "functionalDataPerBlock")

# directory with MRI data in the analysis space (samespace)
dirs$samespace_dir <- here("data", "mri", "processed", "samespace")

# base directory for RSA
dirs$rsa_dir <- here("data", "mri", "rsa")

# directories for RSA correlation matrices
dirs$rsa_roi_corr_mat_dirs <- here("data", "mri", "rsa", "correlation_matrices", rois)

# directories for RSA pattern similarity change
dirs$rsa_roi_ps_change_dirs <- here("data", "mri", "rsa", "pattern_similarity_change", rois)

# directories for the cleaned timeseries data
dirs$rsa_roi_clean_timeseries_dirs <- here("data", "mri", "rsa", "clean_roi_timeseries", rois)

# directories for the relevant volumes of each ROI
dirs$rsa_roi_rel_vol_dirs <- here("data", "mri", "rsa", "relevant_roi_volumes", rois)

# directory for data on which to run RSA
dirs$rsa_dat_dir <- here("data", "mri", "rsa", "data_for_rsa")

dirs$mask_dir <- here("data", "mri", "processed", "group_masks")

# base directory for RSA searchlights
dirs$searchlight <- here("data", "mri", "rsa", "searchlight")

# data for final analysis (this will be shared)
dirs$data4analysis <- here("data_for_analysis")

# directory for source data for nature communications
dirs$source_dat_dir <-here("source_data_natcomms")
if(dir.exists(dirs$source_dat_dir)){unlink(dirs$source_dat_dir)} # delete if already existing to make sure to start with empty directory

# directory to save figures to
fig_dir = here("figures")
if(!dir.exists(fig_dir)){dir.create(fig_dir)}

# create all the directories
dirs_created <- lapply(unlist(dirs), function(x) if(!dir.exists(x)) dir.create(x, recursive=TRUE))
```