---
author: "Jacob Bellmund"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

# Prepare (f)MRI Data
## Regions of Interest
We want to run ROI-based representational similarity analyses to test how the hippocampal-entorhinal system represents the learned temporal relationships. We will focus our analysis on the anterior hippocampus (aHPC) and anterolateral entorhinal cortex (alEC). 

>Our previous work demonstrates representations reflecting the temporal relations of events from one sequence in the anterior hippocampus(21) and the anterior-lateral entorhinal cortex(27). More generally, these regions have been implicated in temporal coding and memory (for review, see(10)). Further, the hippocampus has been linked to inferential reasoning and generalization(46,48,49,51,53). We thus focused our analyses on these regions. 

To define participant-specific ROI masks, we want to combine masks from the individual Freesurfer parcellations of the HPC and EC with masks dividing the subregions of the HPC and EC. Specifically, these are based on the Harvard-Oxford atlas distributed with FSL for the hippocampus and on the EC subregion masks from [Navarro Schröder et al. (eLife, 2015)](http://elifesciences.org/content/4/e06738).

We use the Freesurfer segmentation as run by Lorena Deuker. This was obtained via the recon-all command. All masks will be co-registered to the analysis space, which is the space of the wholebrain functional images. This relies on the FSL transformation matrices and warp files obtained from running FEAT on the wholebrain functional images.

### FreeSurfer Masks {-}

>Region of interest (ROI) masks were based on participant-specific FreeSurfer segmentations (version 6.0.0-2), which yielded masks for the entire hippocampus and entorhinal cortex. These were co-registered to participants’ functional space. 

#### Create Masks from Parcellation {-}
First, we want to create nifti masks from the FreeSurfer parcellation. These are saved in the space of the participant's highres structural space.

```{r, eval = run_prep_rois}

# create freesurfer subfolder for in each ROI folder
invisible(lapply(file.path(dirs$rois_fs_dirs, "freesurfer_highres_space"),
                 function(x) if(!dir.exists(x)) dir.create(x)))

for (sub_id in subjects){
    
  # get the name of the ROIs(fs = from freesurfer, hs = highres space)
  out_roi_fn <- file.path(dirs$rois_fs_dirs, "freesurfer_highres_space", 
                          sprintf("P%s_%s_fs_hs.nii.gz", sub_id, rois_fs))

  # load the structural image
  highres_fn <- here("data", "mri", "processed", "wholebrain", 
                     paste0("VIRTEM_P", sub_id, ".feat"), "reg", "highres.nii.gz")
  highres_nii <- readNIfTI2(highres_fn)
        
  # load the freesurfer output
  freesurf_fn <- here("data", "mri", "freesurfer", 
                      paste0("VIRTEM_P", sub_id), "mri", "aparc+aseg-in-rawavg.nii")
  aparc_nii <- readNIfTI2(freesurf_fn)

  for (i_roi in 1:length(rois_fs)){
    
    # initialize all mask voxels as zeros
    mask_idx <- array(0, dim(img_data(highres_nii)))
    
    # set voxels of the ROI to one            
    mask_idx[img_data(aparc_nii) %in% labels_fs[[i_roi]]] <- 1
    
    # create nifti based on the Freesurfer image
    roi_nii <- highres_nii
    img_data(roi_nii) <- mask_idx
    
    # write the nifti to file
    writenii(nim = roi_nii, filename = out_roi_fn[i_roi])
    
    # find coordinates for sensible slices to plot (most frequent value in each dim)
    coords <- arrayInd(which(as.logical(mask_idx)), dim(mask_idx))
    coords <- c(min(statip::mfv(coords[,1])), 
                min(statip::mfv(coords[,2])), 
                min(statip::mfv(coords[,3])))
    
    # save a diagnostic plot as a PDF
    fname <- tools::file_path_sans_ext(out_roi_fn[i_roi], compression = "TRUE")
    pdf(paste0(fname,".pdf"))
    ortho2(robust_window(highres_nii, probs = c(0, 0.999)), y = roi_nii,
           col.y = roi_colors_fs[i_roi], xyz = coords,
           text = sprintf("P%s: %s", sub_id, rois_fs[i_roi]),
           crosshairs = FALSE)
    invisible(dev.off())
  } 
}
```


To get an overview of the ROIs that we just created, we create a merged PDF of the diagnostic plots. We do this for each ROI separately. The files are saved to the specific ROI folders.

```{r, eval = run_prep_rois}
# create a PDF of diagnostic plots for each ROI
for (i_roi in 1:length(rois_fs)){
  
  # find all the PDF files of this ROI
  fnames <- tools::file_path_sans_ext(
  sprintf("P%s_%s_fs_hs.nii.gz", subjects, rois_fs[i_roi]), compression = "TRUE")
  fnames <- file.path(dirs$rois_fs_dirs[i_roi], "freesurfer_highres_space",
                      paste0(fnames, ".pdf"))
  
  # merge PDFs together
  merged_pdf <- here("data", "mri", "rois", rois_fs[i_roi], 
               paste0(rois_fs[i_roi], "_freesurfer_highres_space.pdf"))
  cmd = paste("pdfunite", paste(fnames, collapse = " "), merged_pdf)
  system(cmd)
}
```


#### Coregister FreeSurfer masks to functional space {-}

Next we coregister the masks from FreeSurfer from the highres structural space to the functional analysis space. For this we will use FSL flirt based on the registration files obtained during preprocessing of the wholebrain functional images. Further we will mask them with our partial field of view mask based on the sequence used during the picture viewing tasks.

To make sure everything went well, we create diagnostic image for every subject, which we collect in a PDF for visual inspection. 

```{r, eval = run_prep_rois}

# create samespace freesurfer subfolder for in each ROI folder
invisible(lapply(file.path(dirs$rois_fs_dirs, "freesurfer_samespace"),
                 function(x) if(!dir.exists(x)) dir.create(x)))

# name of transformation matrix file to move from highres to functional space
highres2func <- here("data", "mri", "processed", "wholebrain", 
                     paste0("VIRTEM_P", subjects, ".feat"),
                     "reg", "highres2example_func.mat")

# use the mean EPI of wholebrain image as a reference
mean_epi <- here("data", "mri", "processed", "samespace", paste0("VIRTEM_P", subjects),
                 paste0("VIRTEM_P", subjects, "_wholebrain.nii.gz"))

# functional mask (picture viewing tasks not scanned with wholebrain coverage)
brain_mask_ss <- here("data", "mri", "processed", "samespace", 
                      paste0("VIRTEM_P", subjects),
                      paste0("VIRTEM_P", subjects, "_common_mask.nii.gz"))

for (i_roi in 1:length(rois_fs)){
  
  # define the file names of ROI files in highres space
  roi_hs <- file.path(dirs$rois_fs_dirs[i_roi], "freesurfer_highres_space", 
                        sprintf("P%s_%s_fs_hs.nii.gz", 
                                subjects, rois_fs[i_roi]))
  
  # define output files in samespace (ss) = analysis space based on the wholebrain EPI
  roi_ss <- file.path(dirs$rois_fs_dirs[i_roi], "freesurfer_samespace", 
                      sprintf("P%s_%s_fs_ss.nii.gz", subjects, rois_fs[i_roi]))
  
  # define file names for masked ROI files
  roi_ss_masked <- file.path(dirs$rois_fs_dirs[i_roi], "freesurfer_samespace", 
                      sprintf("P%s_%s_fs_ss_masked.nii.gz", 
                              subjects, rois_fs[i_roi]))

  
  # apply FSL flirt to move ROI from highres to wholebrain functional space
  out <- mapply(flirt_apply, infile = roi_hs, reffile = mean_epi, 
                initmat = highres2func, outfile = roi_ss,
                verbose = FALSE, retimg = FALSE)
  
  # mask to make sure to reduce to partial field of view 
  invisible(mapply(fsl_maths, file = roi_ss, opts = sprintf("-min %s", brain_mask_ss), outfile = roi_ss_masked,
                   verbose = FALSE, retimg = FALSE))
}
```

```{r eval = run_prep_rois}
# Diagnostic plotting
for (i_roi in 1:length(rois_fs)){
  
  # define file names for binarized masks
  roi_to_plot <- file.path(dirs$rois_fs_dirs[i_roi], "freesurfer_samespace", 
                           sprintf("P%s_%s_fs_ss_masked.nii.gz", 
                                   subjects, rois_fs[i_roi]))
  
  fnames <- vector(mode = "character", length(subjects))
  for (i_sub in 1:length(subjects)){
    
    # load nifti of mean epi and the roi mask
    mean_epi_nii <- readNIfTI2(mean_epi[i_sub])
    roi_nii <- readNIfTI2(roi_to_plot[i_sub])
    
    # find coordinates for sensible slices to plot
    coords <- arrayInd(which(as.logical(img_data(roi_nii))), dim(img_data(roi_nii)))
    coords <- c(min(statip::mfv(coords[,1])), 
                min(statip::mfv(coords[,2])), 
                min(statip::mfv(coords[,3])))
        
    # save a diagnostic plot as a PDF
    fnames[i_sub] <- paste0(tools::file_path_sans_ext(roi_to_plot[i_sub], compression = "TRUE"),".pdf")
    pdf(fnames[i_sub])
    ortho2(robust_window(mean_epi_nii, probs = c(0, 0.999)), y = roi_nii,
           col.y = roi_colors_fs[i_roi], xyz = coords,
           text = sprintf("P%s: %s\nmasked", subjects[i_sub], rois_fs[i_roi]),
           crosshairs = FALSE)
    invisible(dev.off())
  }
  
  # merge PDFs together
  merged_pdf <- here("data", "mri", "rois", rois_fs[i_roi], 
               paste0(rois_fs[i_roi], "_freesurfer_samespace_masked.pdf"))
  #pdf_combine(fnames, output = merged_pdf) # pdftools doesn't work on remote linux?
  cmd = paste("pdfunite", paste(fnames, collapse = " "), merged_pdf)
  system(cmd)
}
```