---
author: "Jacob Bellmund"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

### Subregion masks (MNI) {-}

>We defined anterior hippocampus using the Harvard-Oxford atlas mask (thresholded at 50% probability), selecting all voxels anterior to MNI y=-21 based on Poppenk et al. (2013). The resulting anterior hippocampus mask was also co-registered to participants’ functional space and intersected with the participant-specific hippocampal mask from FreeSurfer. The mask for the anterior-lateral entorhinal cortex was based on Navarro Schröder et al. (2015). It was co-registered to participants’ functional space and intersected with the entorhinal cortex mask from FreeSurfer.

Above, we generated masks of the Hippocampus and Entorhinal Cortex based on the Freesurfer segmentation of each participant's structural image. Our hypotheses concern the anterior portion of the hippocampus and the anterior-lateral entorhinal subregion, specifically.  Hence, we want to split the participant-specific masks. 

* For the hippocampus, we will start out with the probabilistic masks based on the Harvard-Oxford atlas delivered with FSL. Masks for the right and left hippocampus  will be combined, thresholded at 50% probability and split in an anterior and posterior section. We will define the anterior hippocampus as all voxels anterior to MNI y = -21 based on [Poppenk et al. (TiCS, 2013)](https://www.sciencedirect.com/science/article/pii/S1364661313000673). They "propose that foci at or anterior to y = -21 mm in MNI space (y = -20 mm in Talairach space) may be regarded as falling in the aHPC, as this coordinate incorporates the uncal apex in the MNI152 template and current neuroanatomical atlases".
* We will define the anterolateral entorhinal cortex based on [Navarro Schröder et al. (eLife, 2015)](http://elifesciences.org/content/4/e06738). Specifically, we will use a mask for the dominant mode of connectivity change within the EC, which is shown in [Figure 2](https://elifesciences.org/articles/06738/figures#fig2) of the original paper. 

The code below assumes that masks for the left and right hippocampus were extracted from the Harvard-Oxford atlas (MNI space 1mm). Likewise, it assumes the EC subregion masks (MNI space 1mm, obtained from [Tobias Navarro Schröder](https://www.researchgate.net/profile/Tobias_Navarro_Schroeder) to be included in the folder. After the hippocampus masks are combined, thresholded, and split, the masks will be co-registered to the functional analysis space.

```{r, eval = run_prep_rois}

# define the name of the ROIs in MNI space
rois_hpc_fnames <- here("data", "mri", "rois", "mni_masks", paste0(rois_hpc,".nii.gz"))

# combine atlas mask across hemispheres and split into anterior & posterior if not done
atlas_fnames <- here("data", "mri", "rois", "mni_masks", 
                   c("harvardoxford-subcortical_prob_Left_Hippocampus.nii.gz",
                     "harvardoxford-subcortical_prob_Right_Hippocampus.nii.gz",
                     "harvardoxford-hpc_lr.nii.gz",
                     "harvardoxford-hpc_lr_tresh50_bin.nii.gz"))
# combine the left and right HPC mask from the Harvard-Oxford atlas
fsl_add(atlas_fnames[1], atlas_fnames[2],outfile = atlas_fnames[3],
       retimg = FALSE, verbose = FALSE)

# threshold at 50 percent probability
fsl_thresh(atlas_fnames[3], outfile = atlas_fnames[4], thresh = 50, opts = "-bin",
           retimg = FALSE, verbose = FALSE)

# create anterior hippocampus ROI by creating a volume with all voxels anterior
# to MNI y=-21 (y=105 in matrix coordinates) and mask this with the binary HPC ROI
opts = sprintf("-mul 0 -add 1 -roi 0 182 105 -1 0 182 0 1 -mas %s", atlas_fnames[4])
fsl_maths(file = atlas_fnames[4], outfile = rois_hpc_fnames[1], opts = opts,
          retimg = FALSE, verbose = FALSE)
```

Here are some diagnostic plots for the ROIs that we will use (in MNI 1mm space).

```{r, eval = run_prep_rois}
# EC ROIs
rois_ec_fnames <- here("data", "mri", "rois", "mni_masks", paste0(rois_ec,".nii.gz"))

# file names for mni ROIs
rois_mni_fnames <- c(rois_hpc_fnames, rois_ec_fnames)

# load mni 1mm (here we use the ch2better template from MRIcron)
#mni_nii <- here("data", "mri", "rois", "mni_masks", "MNI152_T1_1mm_brain.nii.gz") #%>% readNIfTI2()
mni_nii <- here("data", "mri", "rois", "mni_masks", "ch2better_mni1mm.nii.gz")
# register to MNI 1mm space if needed
if(!file.exists(mni_nii)){
  flirt(infile = here("data", "mri", "rois", "mni_masks", "ch2better.nii.gz"),
        reffile = here("data", "mri", "rois", "mni_masks", "MNI152_T1_1mm_brain.nii.gz"),
        outfile = mni_nii, omat = "ch2better_to_mni1mm")
}
mni_nii <- readNIfTI2(mni_nii)

for (i_roi in 1:length(rois_mni)){
  
  # load roi mask nifti
  roi_nii <- readNIfTI2(rois_mni_fnames[i_roi])
  
  # find coordinates for sensible slices to plot
  coords <- arrayInd(which(as.logical(img_data(roi_nii))), dim(img_data(roi_nii)))
  coords <- c(min(statip::mfv(coords[,1])), 
              min(statip::mfv(coords[,2])), 
              min(statip::mfv(coords[,3])))
    
  # create a diagnostic plot and save as a PDF
  fname <- paste0(tools::file_path_sans_ext(rois_mni_fnames[i_roi], compression = "TRUE"),".pdf")
  ortho2(robust_window(mni_nii, probs = c(0, 0.999)), y = roi_nii, 
         col.y = roi_colors[i_roi], xyz = coords,
         text = rois_mni[i_roi], crosshairs = FALSE)
  dev.copy(pdf, fname)
  invisible(dev.off())
}
```

The resulting ROI masks can now be coregistered from MNI 1mm space to the analysis space of the wholebrain functional sequence. Finally, they are thresholded at a probability of 0.5.

```{r, eval = run_prep_rois}

# name of transformation matrix file to move from highres to functional space
standard2func <- here("data", "mri", "processed", "wholebrain", 
                     paste0("VIRTEM_P", subjects, ".feat"),
                     "reg", "standard2example_func.mat")

# use the mean EPI of wholebrain image as a reference
mean_epi <- here("data", "mri", "processed", "samespace", paste0("VIRTEM_P", subjects),
                 paste0("VIRTEM_P", subjects, "_wholebrain.nii.gz"))

for (i_roi in 1:length(rois_mni)){
  
  #file name of ROI file in standard MNI space
  roi_mni <- rois_mni_fnames[i_roi]
  
  # define output files in samespace (ss) = analysis space based on the wholebrain EPI
  roi_ss <- file.path(dirs$rois_mni_ss_dirs[i_roi],
                      sprintf("P%s_%s_ss.nii.gz", subjects, rois_mni[i_roi]))

  # apply FSL flirt to move ROI from standard to wholebrain functional space
  invisible(mapply(flirt_apply, infile = roi_mni, reffile = mean_epi, 
                   initmat = standard2func, outfile = roi_ss,
                   verbose = FALSE, retimg = FALSE))
  
  # use fslmaths to binarize the masked ROIs using a threshold of 0.5
  out <- mapply(fsl_thresh, file = roi_ss, outfile = roi_ss, thresh = 0.5, opts = "-bin",
                verbose = FALSE, retimg = FALSE)
}
```
