---
author: "Jacob Bellmund"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

## Combine behavioral and fMRI data for RSA

To create input for the RSA the next step is to combine the similarity change data with the behavioral data, so that we can do meaningful analyses. First the behavioral data is loaded and brought into the same pair-wise format as the similarity change data. Both datasets are combined for each subject and then written to disk.


```{r}

for (i_sub in subjects){
  
  # load the behavioral data
  fn <- file.path(dirs$timeline_dat_dir, sprintf("%s_behavior_tbl_timeline.txt", i_sub))
  col_types_list <- cols_only(sub_id = col_character(), day = col_factor(), 
                              event = col_integer(), pic = col_integer(),
                              virtual_time = col_double(), real_time = col_double(),
                              memory_time = col_double(), memory_order = col_double(), 
                              sorted_day = col_integer())
  beh_dat <- read_csv(fn, col_types = col_types_list)
  
  # sort behavioral data according to picture identity
  beh_dat_ordered <- beh_dat[order(beh_dat$pic),]
  
  # find the order of comparisons
  pairs <- which(upper.tri(matrix(nrow = 20, ncol = 20)), arr.ind=TRUE)
  
  # extract the data for the first and second picture in each pair from the behavioral data
  pic1_dat <- beh_dat_ordered[pairs[,1],]
  pic2_dat <- beh_dat_ordered[pairs[,2],]
  colnames(pic1_dat) <- paste0(colnames(pic1_dat), "1")
  colnames(pic2_dat) <- paste0(colnames(pic2_dat), "2")
  
  # combine the two tibbles
  pair_dat <- cbind(pic1_dat, pic2_dat)
  
  # reorder the tibble columns to make a bit more sense
  pair_dat <- pair_dat %>%
    select(sub_id1,
           day1, day2,
           pic1, pic2,
           event1, event2, 
           virtual_time1, virtual_time2, 
           real_time1, real_time2, 
           memory_order1, memory_order2, 
           memory_time1, memory_time2, 
           sorted_day1, sorted_day2) %>%
    rename(sub_id = sub_id1)
  
  rsa_dat <- tibble()
  for (i_roi in 1:length(rois)){
    
    # load the pattern similarity change data for this ROI
    fn <- file.path(dirs$rsa_roi_ps_change_dirs[i_roi], 
                    sprintf("%s_%s_pattern_similarity_change.txt", i_sub, rois[i_roi]))
    ps_change_dat <- read.csv(fn)
    
    # make sure files have the same order
    assertthat::are_equal(c(pair_dat$pic1, pair_dat$pic2), c(ps_change_dat$pic1, ps_change_dat$pic2))
    
    # add column with ROI name
    ps_change_dat <- add_column(ps_change_dat, roi = rois[i_roi])

    # collect the data from this ROI and merge into long data frame
    roi_dat <- cbind(pair_dat, ps_change_dat[,3:6])
    rsa_dat <- rbind(rsa_dat, roi_dat)
  }
  
  # write to file
  fn <- file.path(dirs$rsa_dat_dir, sprintf("%s_data_for_rsa.txt",i_sub))
  write.table(rsa_dat, fn, append = FALSE, sep = ",",
              dec = ".", row.names = FALSE, col.names = TRUE)
}
```


Finally, the datasets are combined across subjects.

```{r}

# set up a dataframe to collect the data
rsa_dat = tibble()

for (i_sub in subjects){

  # load data from CSV
  fn <- file.path(dirs$rsa_dat_dir, sprintf("%s_data_for_rsa.txt",i_sub))
  col_types_list <- cols_only(
        sub_id = col_character(),
        day1 = col_integer(), day2 = col_integer(),
        event1 = col_integer(), event2 = col_integer(),
        pic1 = col_integer(), pic2 = col_integer(),
        virtual_time1 = col_double(), virtual_time2 = col_double(),
        real_time1 = col_double(), real_time2 = col_double(),
        memory_time1 = col_double(), memory_time2 = col_double(), 
        memory_order1 = col_double(), memory_order2 = col_double(), 
        sorted_day1 = col_integer(), sorted_day2 = col_integer(),
        pre_corrs  = col_double(), post_corrs = col_double(),
        ps_change = col_double(), roi = col_factor()
  )
  sub_dat <- as_tibble(read_csv(fn, col_types = col_types_list))
  
  # append to table with data from all subjects
  rsa_dat <- bind_rows(sub_dat, rsa_dat)
}

# sort the data
rsa_dat <- rsa_dat[with(rsa_dat, order(sub_id, day1, day2, event1, event2)),]

# write to file
fn <- file.path(dirs$data4analysis, "rsa_data_rois.txt")
write.table(rsa_dat, fn, append = FALSE, sep = ",",
            dec = ".", row.names = FALSE, col.names = TRUE)
```