---
author: "Jacob Bellmund"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

# RSA in searchlight peak with different threshold

This section is to address a reviewer comment about whether the results of this analysis differ when running it with a different threshold (p<0.001 rather than p<0.01) to define the ROI based on the same-sequence searchlight. The code below is thus copied from the original script and modified to include the different p-threshold.

## Prepare RSA in searchlight peak

To show that overlapping clusters of voxels drive both the within- and across sequence effects, we will run the across-sequence analysis in the cluster of voxels that show the within-day effect. Because these are independent comparisons, we can use the within-day searchlight to define a region of interest for the across-day analysis.

### Create the ROI mask {-}

The first steps are to prepare the ROI mask we want to use. Thus, we need to threshold the ROI from the main analysis in MNI space and move the resulting mask to each participant's functional space for the analysis.

```{r}

threshs <- c("uncorrp001lh")

```

```{r, eval = run_srchlgh_peak_RSA2}

for (i_thresh in threshs){ 
  
  # threshold the searchlight results
  if (i_thresh == "corrp05" | i_thresh == "corrp05lh"){
    in_fn <- file.path(dirs$data4analysis, "searchlight_results", "same_day_vir_time",
                       "same_day_vir_time_randomise_svc_fwhm3_tfce_corrp_tstat1.nii.gz")  
  } else{
    in_fn <- file.path(dirs$data4analysis, "searchlight_results", "same_day_vir_time",
                       "same_day_vir_time_randomise_svc_fwhm3_tfce_p_tstat1.nii.gz")
  }
  
  svc_fn <- file.path(dirs$mask_dir, "svc", "svc_mask.nii.gz")
  bin_fn <- here("data", "mri", "rois", "mni_masks", sprintf("searchlight_same-day_svc_%s.nii.gz", i_thresh))
  fsl_thresh(file = in_fn, outfile = bin_fn, thresh = 0.95, opts = sprintf("-bin -mas %s", svc_fn),
             verbose = FALSE, retimg = FALSE)
  
  # peak cluster is in left hemisphere, so don't include any voxels in right hemisphere
  if (i_thresh == "uncorrp001lh" | i_thresh == "corrp05lh"){
    
    lh_fn <- here("data", "mri", "rois", "mni_masks", "left_hemi.nii.gz")
    fsl_maths(file=bin_fn, outfile = lh_fn, opts = "-mul 0 -add 1 -roi 91 182 0 218 0 182 0 1",
              retimg = FALSE, verbose = FALSE)
    fsl_mul(file=bin_fn, file2=lh_fn, outfile=bin_fn,
            retimg = FALSE, verbose = FALSE)
  }
  
  # let's have a look at the mask we created
  mni_nii <- readNIfTI2(mni_fname("1"))
  roi_nii <- readNIfTI2(bin_fn)
  coords <- c(statip::mfv1(which(roi_nii==1, arr.ind = TRUE)[,1]), 
              statip::mfv1(which(roi_nii==1, arr.ind = TRUE)[,2]),
              statip::mfv1(which(roi_nii==1, arr.ind = TRUE)[,3]))
  ortho2(mni_nii, y = roi_nii, xyz = coords, add.orient = TRUE)
  
  # check the number of voxels in this ROI
  sum(c(roi_nii))
}
```

The resulting ROI mask is now coregistered from MNI 1mm space to the analysis space of the wholebrain functional sequence. Finally, it is thresholded at a probability of 0.5.

```{r, eval = run_srchlgh_peak_RSA2}

for (i_thresh in threshs){ 

  samespace_dir <- here("data", "mri", "rois", "same-day_searchlight", "samespace")
  if (!dir.exists(samespace_dir)){dir.create(samespace_dir, recursive = TRUE)}
  
  # name of transformation matrix file to move from highres to functional space
  standard2func <- here("data", "mri", "processed", "wholebrain", 
                        paste0("VIRTEM_P", subjects, ".feat"),
                        "reg", "standard2example_func.mat")
  
  # use the mean EPI of wholebrain image as a reference
  mean_epi <- here("data", "mri", "processed", "samespace", paste0("VIRTEM_P", subjects),
                   paste0("VIRTEM_P", subjects, "_wholebrain.nii.gz"))
  
  # define output files in samespace (ss) = analysis space based on the wholebrain EPI
  roi_ss <- file.path(samespace_dir, sprintf("P%s_%s_%s_ss.nii.gz", 
                                             subjects, "same-day_searchlight", i_thresh))
  
  # apply FSL flirt to move ROI from standard to wholebrain functional space
  invisible(mapply(flirt_apply, infile = bin_fn, reffile = mean_epi, 
                   initmat = standard2func, outfile = roi_ss,
                   verbose = FALSE, retimg = FALSE))
  
  # use fslmaths to binarize the masked ROIs using a threshold of 0.5
  out <- mapply(fsl_thresh, file = roi_ss, outfile = roi_ss, thresh = 0.5, opts = "-bin",
                verbose = FALSE, retimg = FALSE)
}
```


### Calculating the correlation matrices {-}

The following steps are analogous to the preparation of the functional data in the ROI and searchlight analyses. For the main searchlight analyses we already cleaned the voxel-wise time series and extracted the volumes relevant to RSA. Thus, to calculate the correlation matrices and to calculate pattern similarity changes, we fall back onto the scripts from the main ROI analyses and the searchlight.


```{r, eval = run_srchlgh_peak_RSA2}
for (i_thresh in threshs){ 
   
  # for all 10x10 comparisons we will be averaging all comparisons apart from the diagonal
  # to exclude same_block comparisons
  no_diag <- matrix(data = TRUE, nrow=10, ncol = 10)
  diag(no_diag)<- FALSE
  
  out_dir <- here("data", "mri","rsa","correlation_matrices", "same-day_searchlight")
  if (!dir.exists(out_dir)){dir.create(out_dir, recursive = TRUE)}
  
  for (i_sub in subjects){
    
    # load the ROI based on the searchlight
    roi_fn <-file.path(samespace_dir, sprintf("P%s_%s_%s_ss.nii.gz", 
                                              i_sub, "same-day_searchlight", i_thresh))
    roi_nii <- readNIfTI(roi_fn, reorient=FALSE)
    
    # array indices of ROI voxels
    roi_vox <- which(roi_nii == 1, arr.ind=TRUE)
    sprintf("%s: %d voxels\n", i_sub, nrow(roi_vox))
    
    for (i_run in 1:n_runs){
      
      # load the relevant functional volumes
      rel_vol_fn <- file.path(dirs$searchlight, "rel_vols_4D", 
                                    sprintf("%s_run%02d_rel_vols.nii.gz", i_sub, i_run))
      func_nii <- readNIfTI(rel_vol_fn, reorient=FALSE)
      
      # get the ROI voxels
      rel_dat <- array(NA, c(n_pics*n_blocks, nrow(roi_vox)));  # images in rows (ROI voxels), voxels in columns
      for (i in 1:nrow(roi_vox)) {   # i <- 1
            
          curr_vox <- func_nii[roi_vox[i,1], roi_vox[i,2], roi_vox[i,3],]
          if (sd(curr_vox) > 0) {rel_dat[,i] <- curr_vox} else { stop("zero variance voxel")}
       }
  
      # data is in repetition (row) by voxel (col) format, so we transpose 
      # to get a voxel x repetition format
      rel_dat <- t(rel_dat)
      
      # calculate correlation matrix (trial by trial) for pre and post run
      cor_mat_trial <- cor(rel_dat, rel_dat)
  
      # initialize condition by condition correlation matrix for pre and post run
      corr_mat <- matrix(nrow = 20, ncol = 20)
      
      # loop over all picture comparisons
      for(i_pic1 in 1:20){
        for(i_pic2 in 1:20){
          
          # extract the current 10x10 correlation matrix
          i1 <- (1+(i_pic1-1)*10):(i_pic1*10)
          i2 <- (1+(i_pic2-1)*10):(i_pic2*10)
          curr_mat <- cor_mat_trial[i1, i2]
  
          # average the correlations while excluding diagonal (same block comparisons)
          corr_mat[i_pic1, i_pic2] <- mean(curr_mat[no_diag])
        }
      }
      
      # save the correlation matrix
      fn <- file.path(out_dir, sprintf("%s_%s_%s_%s_corr_mat.txt", 
                                       i_sub, "same-day_searchlight", i_thresh, runs[i_run]))
      write.table(corr_mat, fn, append = FALSE, sep = ",",
                  dec = ".", row.names = FALSE, col.names = FALSE)
    }
  }
}
```

### Calculate Pattern Similarity Change {-}

In the next step, we calculate how the correlation between patterns change from the first to the second picture viewing task, i.e. through the day learning task. To do this, we load both correlation matrices and reduce them to the upper triangle, excluding the diagonal. Then, the correlations are Fisher Z-transformed and the similarity values from the pre-learning picture viewing task are subtracted from those of the post-learning picture viewing task to isolate pattern similarity change. The correlations and their changes are then saved together with information about the pictures that were compared.

This part is based on the corresponding script from the main ROI analyses.
```{r, run_srchlgh_peak_RSA2}

in_dir <- here("data", "mri","rsa","correlation_matrices", "same-day_searchlight")
out_dir <- here("data", "mri","rsa","pattern_similarity_change", "same-day_searchlight")
if (!dir.exists(out_dir)){dir.create(out_dir, recursive = TRUE)}

for (i_thresh in threshs){ 
  
  for(i_sub in subjects){
  
    # load pre correlation matrix
    fn <- file.path(in_dir, sprintf("%s_%s_%s_pre_corr_mat.txt", 
                                    i_sub, "same-day_searchlight", i_thresh))
    corr_mat_pre <- read.table(fn, sep = ",", dec = ".")    
    
    # load post correlation matrix
    fn <- file.path(in_dir, sprintf("%s_%s_%s_post_corr_mat.txt", 
                                    i_sub, "same-day_searchlight", i_thresh))
    corr_mat_post <- read.table(fn, sep = ",", dec = ".")
    
    # reduce to upper triangle of correlations (without diagonal)
    pre_corrs <- corr_mat_pre[upper.tri(corr_mat_pre, diag = FALSE)]
    post_corrs <- corr_mat_post[upper.tri(corr_mat_post, diag = FALSE)]
    
    # create a tibble with the correlations 
    pics <- which(upper.tri(corr_mat_post), arr.ind=TRUE)
    corr_df <- tibble(pic1 = pics[,1], pic2 = pics[,2], pre_corrs, post_corrs)
    
    # Fisher Z transform correlations and calculate pattern similarity change
    # by subtracting the pre-correlations from the post-correlations
    corr_df$ps_change <- FisherZ(corr_df$post_corrs) - FisherZ(corr_df$pre_corrs)
    
    # write to file
    fn <- file.path(out_dir, sprintf("%s_%s_%s_pattern_similarity_change.txt", 
                                     i_sub, "same-day_searchlight", i_thresh))
    write.table(corr_df, fn, append = FALSE, sep = ",",
                dec = ".", row.names = FALSE, col.names = TRUE)
  }
}
```


### Combine behavioral and fMRI data for RSA {-}

To create input for the RSA the next step is to combine the similarity change data with the behavioral data, so that we can do meaningful analyses. First the behavioral data is loaded and brought into the same pair-wise format as the similarity change data. Both datasets are combined for each subject and then written to disk.

This part is based on the corresponding script from the main ROI analyses.

```{r, eval = run_srchlgh_peak_RSA2}

in_dir <- here("data", "mri","rsa","pattern_similarity_change", "same-day_searchlight")
out_dir <- here("data", "mri","rsa","data_for_rsa")
if (!dir.exists(out_dir)){dir.create(out_dir, recursive = TRUE)}

for (i_thresh in threshs){ 
  
  for (i_sub in subjects){
    
    # load the behavioral data
    fn <- file.path(dirs$timeline_dat_dir, sprintf("%s_behavior_tbl_timeline.txt", i_sub))
    col_types_list <- cols_only(sub_id = col_character(), day = col_factor(), 
                                event = col_integer(), pic = col_integer(),
                                virtual_time = col_double(), real_time = col_double(),
                                memory_time = col_double(), memory_order = col_double(), 
                                sorted_day = col_integer())
    beh_dat <- read_csv(fn, col_types = col_types_list)
    
    # sort behavioral data according to picture identity
    beh_dat_ordered <- beh_dat[order(beh_dat$pic),]
    
    # find the order of comparisons
    pairs <- which(upper.tri(matrix(nrow = 20, ncol = 20)), arr.ind=TRUE)
    
    # extract the data for the first and second picture in each pair from the behavioral data
    pic1_dat <- beh_dat_ordered[pairs[,1],]
    pic2_dat <- beh_dat_ordered[pairs[,2],]
    colnames(pic1_dat) <- paste0(colnames(pic1_dat), "1")
    colnames(pic2_dat) <- paste0(colnames(pic2_dat), "2")
    
    # combine the two tibbles
    pair_dat <- cbind(pic1_dat, pic2_dat)
    
    # reorder the tibble columns to make a bit more sense
    pair_dat <- pair_dat %>%
      select(sub_id1,
             day1, day2,
             pic1, pic2,
             event1, event2, 
             virtual_time1, virtual_time2, 
             real_time1, real_time2, 
             memory_order1, memory_order2, 
             memory_time1, memory_time2, 
             sorted_day1, sorted_day2) %>%
      rename(sub_id = sub_id1)
    
    rsa_dat <- tibble()
  
    # load the pattern similarity change data for this ROI
    fn <- file.path(in_dir, sprintf("%s_%s_%s_pattern_similarity_change.txt", 
                                    i_sub, "same-day_searchlight", i_thresh))
    ps_change_dat <- read.csv(fn)
    
    # make sure files have the same order
    assertthat::are_equal(c(pair_dat$pic1, pair_dat$pic2), c(ps_change_dat$pic1, ps_change_dat$pic2))
    
    # add column with ROI name
    ps_change_dat <- add_column(ps_change_dat, roi = paste0("same-day_searchlight_", i_thresh))
  
    # collect the data from this ROI and merge into long data frame
    roi_dat <- cbind(pair_dat, ps_change_dat[,3:6])
    rsa_dat <- rbind(rsa_dat, roi_dat)
    
    # write to file
    fn <- file.path(dirs$rsa_dat_dir, sprintf("%s_data_for_rsa_same-day_searchlight_%s.txt", i_sub, i_thresh))
    write.table(rsa_dat, fn, append = FALSE, sep = ",",
                dec = ".", row.names = FALSE, col.names = TRUE)
  }
}
```


Finally, the datasets are combined across subjects.

```{r, eval = run_srchlgh_peak_RSA2}

# set up a dataframe to collect the data
rsa_dat = tibble()
for (i_thresh in threshs){ 
  
  for (i_sub in subjects){
  
    # load data from CSV
    fn <- file.path(out_dir, sprintf("%s_data_for_rsa_same-day_searchlight_%s.txt",i_sub, i_thresh))
    col_types_list <- cols_only(
          sub_id = col_character(),
          day1 = col_integer(), day2 = col_integer(),
          event1 = col_integer(), event2 = col_integer(),
          pic1 = col_integer(), pic2 = col_integer(),
          virtual_time1 = col_double(), virtual_time2 = col_double(),
          real_time1 = col_double(), real_time2 = col_double(),
          memory_time1 = col_double(), memory_time2 = col_double(), 
          memory_order1 = col_double(), memory_order2 = col_double(), 
          sorted_day1 = col_integer(), sorted_day2 = col_integer(),
          pre_corrs  = col_double(), post_corrs = col_double(),
          ps_change = col_double(), roi = col_factor()
    )
    sub_dat <- as_tibble(read_csv(fn, col_types = col_types_list))
    
    # append to table with data from all subjects
    rsa_dat <- bind_rows(sub_dat, rsa_dat)
  }
  
  # sort the data
  rsa_dat <- rsa_dat[with(rsa_dat, order(sub_id, day1, day2, event1, event2)),]
  
  # write to file
  fn <- file.path(dirs$data4analysis, sprintf("rsa_data_in_same-seq_searchlight_peak_%s.txt", i_thresh))
  write.table(rsa_dat, fn, append = FALSE, sep = ",",
              dec = ".", row.names = FALSE, col.names = TRUE)
}
```

## RSA

We use permutation-based linear model to analyze the data with a summary statistics approach.

```{r}

# define function that calculates z-value for permutation
lm_perm_jb2 <- function(in_dat = df, lm_formula = lm_formula, nsim = 1000){
  
  # run the model for original data and store observed t-values
  lm_fit <- lm(formula = lm_formula, data=in_dat)
  obs <- coef(summary(lm_fit))[,"t value"]
  
  # extract the dependent variable from the formula
  dv <- str_extract(lm_formula, "[^~]+")
  dv <- str_replace_all(dv, fixed(" "), "")
  if(!(dv %in% colnames(in_dat))){stop("Cannot find dependent variable in input data");}
  
  # initialize df for permutation
  data_perm <- in_dat
  
  # set aside space for results
  res <- matrix(nrow = nsim, ncol = length(obs))
  
  for (i in 1:nsim) {
    
    # scramble response value
    perm <- sample(nrow(in_dat))
    dv_dat <- in_dat[dv]
    data_perm[dv] <- dv_dat[perm,]

    # compute linear model and store the t-value of predictor
    lm_fit_perm <- lm(formula = lm_formula, data=data_perm)
    res[i,] <- coef(summary(lm_fit_perm))[,"t value"]
    }
  
  # append the observed value to the list of results
  res <- rbind(res,obs)
  
  # calculate p-value for each coefficient and transform to z
  p <- rep(0,length(obs))
  z <- rep(0,length(obs))
  for (i_coef in 1:length(obs)){
    p[i_coef] <- sum(res[,i_coef] <= obs[i_coef])/nrow(res)
    z[i_coef] <- -1*qnorm(1-p[i_coef])
    }
  return(z)
}
```
 
Now run the linear model for each participant and do group-level stats.

```{r}

for (i_thresh in threshs){ 
  
  # load the data
  col_types_list <- cols_only(
          sub_id = col_character(),
          day1 = col_integer(), day2 = col_integer(),
          virtual_time1 = col_double(), virtual_time2 = col_double(),
          ps_change = col_double(), roi = col_factor())
  fn <- file.path(dirs$data4analysis, sprintf("rsa_data_in_same-seq_searchlight_peak_%s.txt", i_thresh))
  rsa_dat_within_cluster <- as_tibble(read_csv(fn, col_types = col_types_list))
  
  
  ## PREAPARE RSA
  # create predictors for RSA
  rsa_dat_within_cluster <- rsa_dat_within_cluster %>%
    mutate(
      
      # pair of events from say or different day (dv = deviation code)
      same_day = day1 == day2,
      same_day_dv = plyr::mapvalues(same_day, from = c(FALSE, TRUE), to = c(-1, 1)),
      
      # absolute difference in time metrics
      vir_time_diff = abs(virtual_time1 - virtual_time2)) %>%
    
    # z-score the time metric predictors (within each subject)
    group_by(sub_id) %>%
    mutate_at(c("vir_time_diff"), scale, scale = TRUE) %>%
    ungroup()

  ##### First-level RSA {-}
  
  set.seed(179) # set seed for reproducibility
  
  # do RSA using linear model and calculate z-score for model fit based on permutations
  rsa_fit_within_cluster2 <- rsa_dat_within_cluster %>% group_by(sub_id, same_day) %>%
            # run the linear model
            do(z = lm_perm_jb2(in_dat = ., lm_formula = "ps_change ~ vir_time_diff", nsim = n_perm)) %>%
    mutate(z = list(setNames(z, c("z_intercept", "z_virtual_time")))) %>%
    unnest_wider(z)
  
  # add group column used for plotting
  rsa_fit_within_cluster2$group <- factor(1)
  
  # add a factor with character labels for within/across days and one to later color control in facets
  rsa_fit_within_cluster2 <- rsa_fit_within_cluster2 %>% 
    mutate(same_day_char = plyr::mapvalues(same_day, 
                                           from = c(0, 1), 
                                           to = c("across days", "within days"), 
                                           warn_missing = FALSE),
           same_day_char = factor(same_day_char, levels = c("within days", "across days")))
  
  # run a group-level t-test on the different sequence RSA fits in the within-searchlight cluster
  stats <- rsa_fit_within_cluster2 %>% 
    filter(same_day == FALSE) %>% 
    select(z_virtual_time) %>% 
    paired_t_perm_jb (., n_perm = n_perm)
  
  # Cohen's d with Hedges' correction for one sample using non-central t-distribution for CI
  d<-cohen.d(d=(rsa_fit_within_cluster2 %>% filter(same_day == FALSE))$z_virtual_time, 
             f=NA, paired=TRUE, hedges.correction=TRUE, noncentral=TRUE)
  stats$d <- d$estimate
  stats$dCI_low <- d$conf.int[[1]]
  stats$dCI_high <- d$conf.int[[2]]
  
  # print results
  huxtable(stats) %>% theme_article()
}

```

Summary Statistics: t-test against 0 for virtual time across sequences in within-sequence searchlight peak (thresholded at p<0.001)  
t~`r stats$parameter`~=`r round(stats$statistic,2)`, p=`r format(round(stats$p_perm, 3), nsmall = 3)`, d=`r format(round(stats$d, 2), nsmall = 2)`, 95% CI [`r format(round(stats$dCI_low, 2), nsmall = 2)`, `r format(round(stats$dCI_high, 2), nsmall = 2)`]


Compare the results using the two different thresholds with a permutation-based t-test on the difference.

```{r}

# add column for cluster
rsa_fit_within_cluster2 <- rsa_fit_within_cluster2 %>% 
  mutate(thresh = "v2")
rsa_fit_within_cluster <- rsa_fit_within_cluster %>% 
  mutate(thresh = "v1")

# merge data frames and select across-sequence values
rsa_fit_within_cluster_comp <- rbind(rsa_fit_within_cluster, rsa_fit_within_cluster2) %>%
  filter(same_day == FALSE)

# run permutation-based t-test on the difference between the two sets of results
stats <- rsa_fit_within_cluster_comp %>%
  select(-z_intercept) %>%
  pivot_wider(names_from = thresh, values_from = z_virtual_time) %>%
  mutate(thresh_diff = v1 - v2) %>%
  select(thresh_diff) %>%
  paired_t_perm_jb (., n_perm = n_perm)

# print results
huxtable(stats) %>% theme_article()
```

Summary Statistics: t-test between RSA results in searchlight cluster defined on different thresholds (p<0.01 vs. p<0.001)
t~`r stats$parameter`~=`r round(stats$statistic,2)`, p=`r format(round(stats$p_perm, 3), nsmall = 3)`