---
author: "Jacob L. S. Bellmund"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

# Credit
## List of packages
Here is a list of packages used in the analysis and references to them. It is created using the [grateful package](https://github.com/Pakillo/grateful).  

Unfortunately, not all packages provide all information, so there are a few warning messages. The code below generates a markdown file with references that we read in and print below. 

```{r, results='hide'}
invisible(grateful::cite_packages(all.pkg = FALSE, include.rmd=FALSE, style = "cell",
                                  out.format = "md"))
```

```{r, results = 'asis'}
used_pkgs <-readLines(here("virtem_code", "citations.md"))

ref_line <- which(used_pkgs=="References")
used_pkgs[ref_line] <- "### References {-}"
used_pkgs[ref_line+1] <- ""
paste(used_pkgs, collapse = '\n') %>% cat()
```

## Session Info

Lastly, we run session info, using the version from the devtools package.

```{r}
devtools::session_info()
```